Reinforcement learning for options\\on target volatility funds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}

### 2023-02-10 tex/m2ef/dnpp_referee_response.tex

\documentclass{article}
\usepackage[english]{babel}
\usepackage{xcolor}

\usepackage[normalem]{ulem}
\newcommand\soutpars[1]{\let\helpcmd\sout\parhelp#1\par\relax\relax}
\long\def\parhelp#1\par#2\relax{%
  \helpcmd{#1}\ifx\relax#2\else\par\parhelp#2\relax\fi%
}
\newcommand{\change}[1]{{\color{red} {#1}}}%{{#1}}
\newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}

\begin{document}

\noindent {\bf ``Reinforcement learning for options on target volatility funds''.\smallskip\\ Response note to referee's report.}\\

\noindent
We thank the Referee for her interesting comments and all the appreciated suggestions to improve our work. We add few remarks to their comments in the following sections and attach a revised version of the paper. We believe that these extensions address all the points highlighted in the reviews and we hope that they will be considered sufficient to recommend our work for the publication.\\

{\bf $\bullet$ Answers to referee's report}\\

We list here a detailed list of responses Referee's major remarks. We address the minor remarks directly in the text.

\begin{enumerate}

 \item \textit{[Both in Abtract and Introduction, the presentation of the problem to deal with is not well structured. As a matter of fact, first the TVS is briefly introduces, then options and TVO are only hinted, and lastly the problem is shortly described, without all these topics are well linked each other. I suggest the authors to improve this presentation.]}

 ...

 \item \textit{[The Authors address the investigated continuous stochastic control problem through two different approaches: an exact one providing the analytical solution of the problem, and an approximate one based on RL. As for the previous point, even in this one the connections between the two solution approaches are not entirely intelligible. In other terms, it is not entirely clear how the two approaches relate each other. Is the approximate one necessary given that the exact approach provides an analytical solution to the problem? What is the "utility" to use an approximate approach given the applicability of an exact one? I suggest the Authors to clarify this issue.]}

 ...

 \item \textit{[In Subsection 3.3, three (so-called intuitive) strategies are introduced, $S_A$, $S_B$ and $S_C$. No rationales are provided concerning their choices. I suggest the Author to give some details about this point.]}

 ...

 \item \textit{[In Section 4, the Authors state that ``[t]he first [RL] method is a specific algorithm developed by us to fit the problem [...]''. As know, an important feature of any RL method is its convergence to the (unknown) best policy, if any. The Authors do not spend words about this characteristic of the specific algorithm they developed and applied, so I ask them to give some detail about this topic.]}

 ...

 \item \textit{[In Subsection 5.1, the Author claim that an appropriate choice of the learning rate starting value constitutes ``a good choice in terms of [...] avoiding over-fitting''. As known, when training a FFNN, standard methods for avoiding over-fitting are based on error metrics calculated over validation sets or through $K$-fold validation approaches. So, it is not clear as the Authors can guarantee no over-fitting without exploiting such methods or similar. I suggest the Authors to carefully clarify this issue.]}

 ...

 \end{enumerate}

\end{document}

### 2023-02-10 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Add files via upload

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx,mathtools}
\usepackage{graphicx,tikz}
\usepackage{lipsum}
\usepackage{enumerate} 
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\graphicspath{{graphs/}}

\definecolor{mylightyellow}{rgb}{1,1,.8}
\definecolor{mylightgreen}{rgb}{.8,1,.8}
\definecolor{mydarkred}{RGB}{178,34,34}
\definecolor{mydarkgreen}{RGB}{34,139,34}
\definecolor{mydarkblue}{RGB}{72,61,139}
\definecolor{mydarkyellow}{RGB}{218,165,32}

\definecolor{myblueA}{RGB}{52,41,39}
\definecolor{myblueB}{RGB}{92,81,109}
\definecolor{myblueC}{RGB}{132,121,179}
\definecolor{myblueD}{RGB}{172,161,249}
\definecolor{myblueE}{RGB}{212,201,255}

\usetikzlibrary{arrows,patterns,shapes,snakes,shadows,fit,backgrounds,positioning,decorations.pathmorphing,external}
\tikzstyle{nameusd} = [circle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, minimum height=4em]
\tikzstyle{nameeur} = [circle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, minimum height=4em]
\tikzstyle{collusd} = [rectangle,fill=mydarkblue!10, inner sep=0.2cm, rounded corners=5mm]
\tikzstyle{colleur} = [rectangle,fill=mydarkred!10, inner sep=0.2cm, rounded corners=5mm]
\tikzstyle{fixto} = [draw, -latex']
\tikzstyle{fixfrom} = [draw, latex'-]
\tikzstyle{floatto} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, -latex']
\tikzstyle{floatfrom} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, latex'-]
\tikzstyle{investor} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{trader} = [circle, draw, top color=white, bottom color=blue!30, draw=blue!50!black!100, drop shadow, minimum height=4em]
\tikzstyle{bank} = [rectangle, draw, top color=white, bottom color=red!20, draw=red!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{market} = [rectangle, draw, top color=white, bottom color=green!20, draw=green!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{yellowbox} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedyellowbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkyellow!25, draw=mydarkyellow!50, drop shadow, rounded corners, text centered]
\tikzstyle{redbox} = [rectangle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedredbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkred!25, draw=mydarkred!50, drop shadow, rounded corners, text centered]
\tikzstyle{bluebox} = [rectangle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedbluebox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkblue!25, draw=mydarkblue!50, drop shadow, rounded corners, text centered]
\tikzstyle{greenbox} = [rectangle, draw, top color=white, bottom color=mydarkgreen!50, draw=mydarkgreen!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedgreenbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkgreen!25, draw=mydarkgreen!50, drop shadow, rounded corners, text centered]

\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}

% \pgfplotstableread{data/total_consumption_example.dat}\swingtotal

\usepackage[]{hyperref}
\hypersetup{
	breaklinks, 
	baseurl = http://,
	pdfborder = 0 0 0,
	pdfpagemode = UseNone, % do not show thumbnails or bookmarks on opening
	pdfstartpage = 1,
	colorlinks=true,breaklinks=true,
	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
	pdftitle = {Reinforcement Learning for Options on Target Volatility Funds},
	pdfsubject = {Reinforcement Learning for Options on Target Volatility Funds}
	pdfkeywords = {Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation, Stochastic Control Problem}
}

\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,maxbibnames=99,sorting=nyt,sortcites=false, natbib=true]{biblatex}
\addbibresource{Bib.bib}

\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}
\setlength\bibitemsep{1.3\itemsep}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

\newcommand{\Eq}[1]{{\begin{equation}{#1}\end{equation}}}

\newcommand{\Exo}[1]{\mathbb{E}\!\left[\,#1\,\right]}
\newcommand{\Ex}[2]{\mathbb{E}_{#1}\!\left[\,#2\,\right]}
\newcommand{\ExG}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal G}_{#1}\right.\right]}
\newcommand{\ExF}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal F}_{#1}\right.\right]}
\newcommand{\ExFT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal F}_{#1}\right]}
\newcommand{\ExGT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal G}_{#1}\right]}
\newcommand{\ExCo}[2]{\mathbb{E}\!\left[\left.\,#1\,\right|\,#2\,\right]}
\newcommand{\ExC}[3]{\mathbb{E}_{#1}\!\left[\left.\,#2\,\right|\,#3\,\right]}
\newcommand{\ExCT}[4]{\mathbb{E}_{#1}^{#2}\!\left[\left.\,#3\,\right|\,#4\,\right]}
\newcommand{\ExT}[3]{\mathbb{E}_{#1}^{#2}\!\left[\,#3\,\right]}
\newcommand{\Exf}[2]{\hat{\mathbb{E}}_{#1}\!\left[\,#2\,\right]}
\newcommand{\ExfT}[3]{\hat{\mathbb{E}}_{#1}^{#2}\!\left[\,#3\,\right]}
\newcommand{\Px}[1]{\mathbb{P}\left\{\,#1\,\right\}}
\newcommand{\PxC}[2]{\mathbb{P}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
\newcommand{\Qx}[1]{\mathbb{Q}\left\{\,#1\,\right\}}
\newcommand{\QxC}[2]{\mathbb{Q}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
\newcommand{\QxCT}[3]{\mathbb{Q}^{#1}\!\left\{\,#2\,\left.\right|\,#3\,\right\}}
\newcommand{\QxT}[2]{\mathbb{Q}^{#1}\left\{\,#2\,\right\}}
\newcommand{\Qxf}[1]{\hat{\mathbb{Q}}\left\{\,#1\,\right\}}
\newcommand{\Var}[2]{\operatorname{Var}_{#1}\!\left[\,#2\,\right]}
\newcommand{\Covar}[3]{\operatorname{Covar}_{#1}\!\left[\,#2,#3\,\right]}
\newcommand{\ind}[1]{1_{\{#1\}}}
\newcommand{\clip}[3]{\operatorname{clip}\!\left({#1},{#2},{#3}\right)}
\newcommand{\one}{\ensuremath{\mathbbm{1}}}
\newcommand{\onehalf}{\tfrac{1}{2}}

\newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
\DeclareMathOperator{\Tr}{Tr}
\let\oldnorm\norm   % <-- Store original \norm as \oldnorm
\let\norm\undefined % <-- "Undefine" \norm
\DeclarePairedDelimiter\norm{\lVert}{\rVert}

\newcommand{\rec}{R}
\newcommand{\lgd}{\mbox{L{\tiny GD}}}
\newcommand{\virgolette}[1]{``#1''}
\newcommand{\pvcall}{{\mathcal{C}}}
\newcommand{\pvput}{{\mathcal{P}}}
\newcommand{\rom}[1]{\uppercase\expandafter{\romannumeral #1\relax}}

\usepackage{eurosym}
\newcommand{\euroms}[0]{\mbox{\small \euro}}
\usepackage{cleveref}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[R]{}
\lhead{\footnotesize\slshape Daluiso, Nastasi, Pallavicini, Polo, Optimal Strategies for Target Volatility Contracts}
\chead{}
\rhead{\thepage}
\lfoot{}
\cfoot{}
\rfoot{}
\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Reinforcement Learning for Options on \\Target Volatility Funds}

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Group Milano, \texttt{roberto.daluiso@intesasanpaolo.com}}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A. Milano, \texttt{emanuele.nastasi@marketz.eu}}
	\and Andrea Pallavicini\thanks{Intesa Sanpaolo Group Milano, \texttt{andrea.pallavicini@intesasanpaolo.com}} 
	\and
	Stefano Polo\thanks{X-Numeris S.r.l. Milano, \texttt{sp.stefano.polo@gmail.com}} 
}

\begin{document} % goes here

@@

\begin{abstract}
\footnotesize	
\noindent
In this work we deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a control problem to evaluate the option prices. We derive an analytical solution of the problem in the Black and Scholes (BS) scenario. Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an \textit{a priori} solution is not available. We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.
\end{abstract}
\smallskip
\noindent
\textbf{JEL classification codes:} C63, C45, G13.\\
\textbf{AMS classification codes:} 65C05, 68T07, 91G20.\\
\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Target Volatility, Asset Allocation, Stochastic Optimal Control Problem. 
\newpage
\tableofcontents
 \mbox{}
 \vfill
 \noindent \small The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.

\newpage
\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}

@@

These products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew2011} and \textcite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison2013}. At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.

In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard2013} and \textcite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim2018}. Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners. Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular  \textcite{DiGraziano2012}, \textcite{Grasselli2016}, \textcite{Albeverio2019}, and \textcite{DiPersio2019}.

This paper is the first attempt, to the best of our knowledge, to study the funding costs coming from hedging the risky assets underlying the target volatility strategy. In particular, we consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy in term of the financing costs. Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition. In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \citep{BlackScholes1973} and that the derivative contract is a European-style option.

Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem. More precisely we numerically study the problem in the general case of a local volatility dynamics \citep{Dupire1994,Derman1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \textcite{Schulman2017}. 
\\

@@

A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank. Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.

\section{Target Volatility Strategy}\label{sec:TVS}
As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio. Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested in a TVS. In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value. Clients investing in the fund pay a running fee for the service of the fund manager and their capital is protected.

The fund manager usually buys from a bank an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+ \; ,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+ \; .
\end{equation}
In this way, the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.  The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
We remark that the choices of the manager trading activity are stochastic processes since they will depend on the market evolution. Thus, neither the manager strategy nor the bank financing costs for hedging the option can be written in the fund-bank contract. For this reason, the price of a financial product sold by the bank is adjusted to include any valuation adjustment due to the trading activity.

@@

We consider a fund trading a basket of $n$ risky securities with price process $S_t = (S_t^i, \, i=1,\dots,n)$ funded with a cash account $B_t$ accruing at $r_t$. Any dividend paid by the securities is re-invested in the fund. Here, we assume that the TVS is implemented in continuous time, even if in the practice we can implement the strategy only on a discrete set of dates. We introduce the deflated gain process $\Bar{G}_t^i$ associated with the risky securities as given by
\begin{equation}
    \Bar{G}_t^i \coloneqq \Bar{S}_t^i + \Bar{D}_t^i \; ,
\end{equation}
where we define the deflated price and cumulative dividend processes as 
\begin{equation}
    \bar{S}_t^i \coloneqq \frac{S_t^i}{B_t} \; , \quad \bar{D}^i_t \coloneqq \int_0^t\frac{d\pi_u^i}{B_u} + \int_0^t\frac{d\psi_u^i}{B_u} \; ,
\end{equation}
where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
Since fund managers allocating TVS usually rely on equity assets, here we use the results of \textcite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
\begin{equation}
    \psi_t^i \coloneqq \int_0^t S_u^i\mu_u^i du \; ,
\label{eq:XVA_equity}\end{equation}
where we call $\mu_t^i$ cost of carry, which basically represents the hedging costs for the $i$-th security.

@@

\section{Derivative Pricing}\label{sec:Derivative}
In this section, we analyze derivative contracts linked to the TVS described previously. In particular we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.

@@

Thus the derivative price is given by:
\begin{equation}
    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
\label{eq:recursion2}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right) \; .
\end{equation}

\subsection{Black and Scholes Model}\label{subsec:BS_model}
In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift defined in Equation~\eqref{eq:local_drift} can be written as
\begin{equation}
     \ell_\alpha \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\norm{\alpha\left(t,K\right) \cdot \nu\left(t\right)}} \; ,  
\end{equation}
so that the optimization problem can be solved by looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic functions of time.

>>>>

%June 2019

\documentclass[runningheads]{m2ef}

\usepackage[]{latexsym}
\usepackage{graphicx}

\usepackage{amssymb,amsmath,latexsym,amscd,dsfont,enumerate,bbm,mathabx,mathtools}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
\newcommand{\quotations}[1]{``#1''}
\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
\let\oldnorm\norm   % <-- Store original \norm as \oldnorm
\let\norm\undefined % <-- "Undefine" \norm
\DeclarePairedDelimiter\norm{\lVert}{\rVert}

\usepackage[normalem]{ulem}
\newcommand\soutpars[1]{\let\helpcmd\sout\parhelp#1\par\relax\relax}
\long\def\parhelp#1\par#2\relax{%
  \helpcmd{#1}\ifx\relax#2\else\par\parhelp#2\relax\fi%
}
\newcommand{\change}[1]{{\color{red} {#1}}}%{{#1}}
\newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}
\newcommand{\changeB}[1]{{\color{blue} {#1}}}%{{#1}}
\newcommand{\removeB}[1]{{\color{blue} \soutpars{{#1}}}}%{}

\begin{document}

\pagestyle{myheadings} \setcounter{page}{1}

\mainmatter

\title{Reinforcement learning for options\\on target volatility funds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}}

\titlerunning{Reinforcement learning for options on target volatility funds}

\author{Roberto Daluiso\inst{1}
\and Emanuele Nastasi\inst{2} \and
Andrea Pallavicini\inst{1} \and Stefano Polo\inst{3}}

\authorrunning{R. Daluiso, E. Nastasi, A. Pallavicini, S. Polo}

\institute{Financial Engineering, Intesa Sanpaolo Group,\\
Largo Mattioli 3, 20121 Milan, Italy\\
\email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
Financial Engineering, Marketz S.p.A.,\\
Corso Europa, 2, 20122 Milan, Italy\\
\email{Emanuele.Nastasi@marketz.eu} \and Financial Engineering,
X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
\email{Stefano.Polo@xnumeris.com}}

@@

{\let\thefootnote\relax\footnote{\vskip1mm$\!\!\!\!\!\!\!\!\!\!\!\!\!${\fbox{{\parbox[c][1.0cm]{12cm}{\center{\textsc{\vskip-4.5mm \textbf{Mathematical Methods in Economics and Finance}}} \textbf{-- m$\mathbf{^2}$ef} \\ Vol. $\ldots$, No. $\ldots$, $\ldots$\\}}}}}} \setcounter{footnote}{0}

\begin{abstract}
We deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a control problem to evaluate the option prices. We derive an analytical solution of the problem in the Black and Scholes (BS) scenario. Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an {\it a priori} solution is not available. We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.

\smallskip

\noindent {\bf Keywords.} Reinforcement learning, target volatility strategy, hedging costs.

\smallskip

\noindent {\bf M.S.C. classification.} \textsc{\small 65C05, 68T07, 91G20.}

\noindent {\bf J.E.L. classification.} \textsc{\small C63, C45, G13.}
\end{abstract}

\bigskip

@@

These products were initially offered in the Asian markets, see for instance the reports of \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}. At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.

In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}. Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners. Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.

This paper is the first attempt, to the best of our knowledge, to study the funding costs coming from hedging the risky assets underlying the target volatility strategy. In particular, we consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy in term of the financing costs. Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition. In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.

Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem. More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}. 
\\

@@

A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank. Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.
\\

The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.

\section{Target volatility strategy}\label{sec:TVS}
As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio. Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested in a TVS. In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value. Clients investing in the fund pay a running fee for the service of the fund manager and their capital is protected.

The fund manager usually buys from a bank an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+ \; ,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+ \; .
\end{equation}

In this way, the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.  The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
We remark that the choices of the manager trading activity are stochastic processes since they will depend on the market evolution. Thus, neither the manager strategy nor the bank financing costs for hedging the option can be written in the fund-bank contract. For this reason, the price of a financial product sold by the bank is adjusted to include any valuation adjustment due to the trading activity.

@@

We consider a fund trading a basket of $n$ risky securities with price process $S_t = (S_t^i, \, i=1,\dots,n)$ funded with a cash account $B_t$ accruing at $r_t$. Any dividend paid by the securities is re-invested in the fund. Here, we assume that the TVS is implemented in continuous time, even if in the practice we can implement the strategy only on a discrete set of dates. We introduce the deflated gain process $\Bar{G}_t^i$ associated with the risky securities as given by
\begin{equation}
    \Bar{G}_t^i \coloneqq \Bar{S}_t^i + \Bar{D}_t^i \; ,
\end{equation}
where we define the deflated price and cumulative dividend processes as 
\begin{equation}
    \bar{S}_t^i \coloneqq \frac{S_t^i}{B_t} \; , \quad \bar{D}^i_t \coloneqq \int_0^t\frac{d\pi_u^i}{B_u} + \int_0^t\frac{d\psi_u^i}{B_u} \; ,
\end{equation}
where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
\begin{equation}
    \psi_t^i \coloneqq \int_0^t S_u^i\mu_u^i du \; ,
\label{eq:XVA_equity}\end{equation}
where we call $\mu_t^i$ cost of carry, which basically represents the hedging costs for the $i$-th security.

@@

\section{Derivative pricing}\label{sec:Derivative}
In this section, we analyze derivative contracts linked to the TVS described previously. In particular we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.

@@

Thus the derivative price is given by:
\begin{equation}
\begin{aligned}
	    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)= \\ 
	     \mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right] \; ,
\end{aligned}
\label{eq:recursion2}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right) \; .
\end{equation}

\subsection{Black and Scholes model}\label{subsec:BS_model}
In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift defined in Equation~\eqref{eq:local_drift} can be written as
\begin{equation}
     \ell_\alpha \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\norm{\alpha\left(t,K\right) \cdot \nu\left(t\right)}} \; ,  
\end{equation}
so that the optimization problem can be solved by looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic functions of time.
